---
title: "Explanation_FARSFunctions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Explanation_FARSFunctions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(testFars)
```


### BASICS

This package is intended to facilitate the use of information on fatality available via the website of the [Fatality Analysis Reporting System (FARS)](https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars). It allows then plotting the location of fatalities per year and state as wished by the user.

The package consists of five relatively simple functions, three of them are used internally by the two functions that are available for the general user. The use of these two functions is then explained in this simple tutorial. 

##### fars_summarize_years

This is the first function which produces a table with how many fatalities occurred in (a) specific year (s) provided as parameter by the user. The function summarizes the information by month.

```{r,message=TRUE}

## Having tried several options it only works when it calls the data from the directory...
data_dir <- "/home/ruth/Dokumente/Learning/RProgramming/testFars/inst/extdata"
setwd(data_dir)


p1 <- system.file("extdata", "accident_2013.csv.bz2", package = "testFars")
p2 <- system.file("extdata", "accident_2014.csv.bz2", package = "testFars")
p3 <- system.file("extdata", "accident_2015.csv.bz2", package = "testFars")
# with one year
accident_2013.csv.bz2 <- sample_read(p1)
#print(head(accident_2013.csv.bz2))

fars_summarize_years(2013)

# woth more than one year 
fars_summarize_years(c(2013,2014,2015))

```

#### fars_map_state

The second function *fars_map_state* accepts two parameters, the first one is an integer number which specifies a state of the United States of America and the second one a year (only one). This function produces a map with the location of the fatalities in that year, in that specified state. For example if we wanted to 

```{r,message=TRUE}
setwd(data_dir)
options(width=300)
# Plotting the first state Alabama and three different years of fatalities
par(mfrow=c(1,3))
fars_map_state(1,2013)
mtext(text=2013,side=3,outer=FALSE)
fars_map_state(1,2014)
mtext(text=2014,side=3,outer=FALSE)
fars_map_state(1,2015)
mtext(text=2015,side=3,outer=FALSE)

# The map of Minnesota 
# Note that fatalities concentrate of course in the area of the twin cities where the largest populations are
fars_map_state(27,2015)



